% \documentclass[compacto,10pt]{aleph-notas}
\documentclass[compacto,10pt,comentarios]{aleph-notas}

% -- Paquetes adicionales
\usepackage{enumitem}
\usepackage{aleph-comandos}
\usepackage{parskip}
\usepackage{graphicx}
\usepackage{xfrac}
\usepackage{tikz}
\usepackage{etoolbox}
\usepackage[framemethod=tikz]{mdframed}
\DeclareFontFamily{U}{skulls}{}
\DeclareFontShape{U}{skulls}{m}{n}{ <-> skull }{}
\newcommand{\skull}{\text{\usefont{U}{skulls}{m}{n}\symbol{'101}}}
\def \ds{\displaystyle}
\def \dfx{\dfrac{d}{dx}}
\DeclareMathOperator{\arccot}{arccot}
\DeclareMathOperator{\arcsec}{arcsec}
\DeclareMathOperator{\arccsc}{arccsc}
\newcommand*\Heq{\ensuremath{\overset{\kern2pt LH}{=}}}

% -- Datos del libro
\institucion{Southwestern College}
\asignatura{MATH 251: Calculus II}
\tema{Differential Equations: Basic Ideas}
\autor{Jesús Pérez Cuarenta}
% \fecha{Fall 2024}

%% --> Logos de las guias
\logouno[4.5cm]{../Images/swc_logo}
\definecolor{colordef}{cmyk}{0.81, 0.62, 0.00, 0.22}

%% -- Solucion para alumnos
\AtBeginEnvironment{proof}{\color{white}}

%% 
% https://www.emathhelp.net/en/calculators/algebra-2/partial-fraction-decomposition-calculator/

\begin{document}

\encabezado

\section*{Differential Equations: Basic Ideas}
\begin{mdframed}
    \center Learning Objectives \\
    \begin{itemize}
        \item Recognize differential equations and verify solutions.
        \item Solve first-order differential equations (general and particular solutions).
    \end{itemize}
\end{mdframed}

A differential equation involves an unknown function $y$ and its derivatives. The unknown in a differential equation is not a number (as in algebraic equation), but rather a \textit{function}.
\begin{itemize}
    \item Examples of algebraic equations
    \begin{align*}
        & x^2 + x + 1 = \mathrm{e}^{x}, \\
        & \sin(x) = x + 1, \\ 
        & \ln(x) = \mathrm{e}^{(x)} ~ .
    \end{align*}
    \item Examples of differential equations
    \begin{align*}
        & \frac{dy}{dx} + 4y = \cos(x), \\
        & \frac{d^2y}{dx^2} + 16y = 0, \\
        & y'(t) = 0.1y(100 - y) ~ .
    \end{align*}
\end{itemize}
    Consider the differential equation
    $$
        y' = y ~ .
    $$
    The solution is a function that produces itself upon differentiation. One example being the exponential function, $\mathrm{e}^{(x)}$ (up to a constant). Can you come up with another example? In order to omit trivial solutions, we require an initial condition.

    Let's look at another case
    $$
        y' = 2x
    $$
    where solutions (which we call \textbf{general solutions}) are of the form 
    $$
        y = x^{2} + C ~ .
    $$
    On the other hand, if we provide an initial condition, say $(x_0, y_0) = (0, 1)$, it follows that
    $$
        1 = 0^2 + C \implies C = 1
    $$
    so that
    $$
        y = x^{2} + 1 
    $$
    is then a \textbf{particular solution}.

\begin{ejer}
    Show by substitution that the function
    $$
        y(t) = C \mathrm{e}^{2.5t}
    $$
    is a solution of the differential equation
    $$
        y'(t) = 2.5 y(t),
    $$
    where $C$ is an arbitrary constant.
\end{ejer}
\begin{proof}[Solution]
    We differentiate $y(t)$,
    $$
    y'(t) = \frac{d}{dt} C \mathrm{e}^{2.5t} = C \cdot 2.5 \mathrm{e}^{2.5t} = 2.5 \cdot C \mathrm{e}^{2.5t} = 2.5y(t)
    $$
    and we are done.
\end{proof}

\begin{ejer}
    Show by substitution that the function
    $$
        y(t) = 3.2 \mathrm{e}^{2.5t}
    $$
    satisfies the initial value problem
    \begin{align*}
        y'(t) & = 2.5 y(t), \\
        y(0) & = 3.2 ~ .
    \end{align*}
\end{ejer}
\begin{proof}[Solution]
    First we evaluate $y(t)$ at $t=0$ and obtain
    \begin{align*}
        y(0) = 3.2 \mathrm{e}^{2.5 (0)} = 3.2
    \end{align*}
    which implies the initial value is satisfied. Now we differentiate and get
    \begin{align*}
        y'(t) & = \frac{d}{dt} 3.2 \mathrm{e}^{2.5t} \\
        & = 3.2 \cdot 2.5\mathrm{e}^{2.5t} \\
        & = 2.5 \cdot 3.2 \mathrm{e}^{2.5t} \\
        & = 2.5 y(t) .
    \end{align*}
    Therefore, $y(t) = 3.2\mathrm{e}^{2.5t}$ is a particular solution.
\end{proof}

\begin{ejer}
    Find the general solution of the following differential equation
        $$
            y'(t) = 5\cos(t) + 6 \sin(3t) ~ .
        $$
\end{ejer}
\begin{proof}[Solution]
    We integrate both sides of the differential equation with respect to $t$,
    \begin{align*}
        \int y'(t) ~ dt & = \int 5\cos(t) + 6 \sin(3t) ~ dt \\
        \implies \int dy & = \int 5\cos(t) + 6 \sin(3t) ~ dt \\
        \implies y & = 5\sin(t) - 2\cos(3t) + C
    \end{align*}
\end{proof}


\end{document}